\documentclass[14pt,letterpaper]{extarticle}
\usepackage[utf8]{inputenc}

% Set margins
% https://tex.stackexchange.com/a/327136
% this should replace use of the 'fullpage' package
\usepackage[
    includehead,
    margin=2.5cm,
    bmargin=2cm,
]{geometry}

\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{verbatim} % This works better than comment for some reason
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{gensymb}
\usepackage{enumitem}
\usepackage{xfrac} % Fractions - https://tex.stackexchange.com/a/230973
\usepackage{xcolor}
% The magic latex drawing thing
\usepackage{tikz}
\usetikzlibrary{shapes}
\usepackage[en-US,showdow]{datetime2}

% Colored hyperlinks:
% https://www.overleaf.com/learn/latex/Hyperlinks
\hypersetup{
    colorlinks=true,
    urlcolor=blue
}

% Setup nice headings
% https://www.overleaf.com/learn/latex/Headers_and_footers
\usepackage{fancyhdr}
\usepackage{titling} % Save title for use in fancyhdr

% Shortcuts
\newcommand*{\bP}{\mathbf{P}}



% Use black tombstone for QED
\renewcommand{\qedsymbol}{$\blacksquare$}

% Important sets
\newcommand*{\C}{\mathbb{C}}
\newcommand*{\R}{\mathbb{R}}
\newcommand*{\Q}{\mathbb{Q}}
\newcommand*{\Z}{\mathbb{Z}}

%
% symbols
%

% name by meaning, not by character
\newcommand{\grad}{\nabla}


% New command for 'partial QED' using white tombstone
\newcommand*{\partialqed}{$\hfill \square$ \\}

\newcommand*{\todo}{{\color{red}\textbf{TODO:}} }

% Increased spaces spacing
\usepackage{setspace} \onehalfspacing

\title{Math 468 - Homework 3}
\author{Nicholas Schlabach (nickninja@arizona.edu)}
\DTMsavedate{titledate}{2026-02-11}
\date{\DTMUsedate{titledate}}

%% Copied from https://tex.stackexchange.com/a/2238
%% Modified by Techcable 9/19/21 to use '[...]' instead of '(...)'
\newenvironment{amatrix}[1]{%
  \left[\begin{array}{@{}*{#1}{c}|c@{}}
}{%
  \end{array}\right]
}

\begin{document}
\pagestyle{fancy}
\fancyhead[L]{Nicholas Schlabach}
\fancyhead[C]{\hspace{5mm}\Big(\thetitle{}\Big)}
\fancyhead[R]{{\DTMsetdatestyle{mmddyy}\DTMusedate{titledate}}, Page \thepage}
\fancyfoot{} % no footer
\maketitle
% maketitle automatically sets pagestyle=plain, override this
\thispagestyle{fancy}


\begin{center}
\begin{enumerate}
\setcounter{enumi}{1}
\item
    \begin{enumerate}
    \item
        This is not necessarily true for infinite markov chains. Consider the random walk on $\mathbb{Z}$ with probability $.9$ to go to the right and $.1$ to the left. All the states form a single communication class, as there is a nonzero probability of going between any two fixed states. It follows that the graph is both irreducible and closed. \\
        Despite this, the drift to the right means there are no recurring states.
    \item
        If $C$ is a finite closed communication class, then $C$ is recurrent.
    \end{enumerate}
\setcounter{enumi}{3}
\item
    First handle the case $0<p<1$: The states $\{1,\dots,d-1\}$ are transient and $\{0,d\}$ are recurrent. Fix a state $0<x<d$, because $T_0$ is absorbing, $P_x(T_0<T_d)$ is equivalent to $P_x(T_0<\infty)$ regardless of $0<x<d$. \\
    Professor says that $P_?(T_0<\infty)$ is called an "absorbption probability." The sum $P_1(T_0<\infty)=\sum_{r=0}^\infty {2r+1 \choose r}q^{r+1}p^r$ does not work because you can not choose $d+7$ moves right then $d+8$ moves left. You would encounter the absorbing state first! \\
    Instead, define the following function $\alpha(j)=P_j(T_0< \infty)$. Clearly $\alpha(0)=1$. \\
    Assuming we start at $0<j<d$, the event $\{T_0<\infty\}$ can be partitioned into $\{T_0<\infty,X_1=j-1\}, \{T_0<\infty,X_1=j+1\}$. \\
    Using the partition theorem we get the following:
    \[ \alpha(j)=P_{j-1}(T_0<\infty)q+P_{j-1}(T_0<\infty)p=\alpha(j-1)q+\alpha(j+1)p \]
    This is a second-order difference equation, analogous to a second-order differential equation $ay''(t)+by'(t)+cy(t)=0$. \\
    Guess: $\alpha(j)=\beta^j$ to get $\beta^j=q\beta^{j-1}+p\beta^{j+1}$. Simplifying gives $0=p\beta^2-\beta+q$.  This gives $\beta_{1,2}=\frac{1\pm \sqrt{1^2-4(pq)}}{2p}$ then we get $\alpha(j)=c_1\beta_1^j+c_2\beta_2^j$ where $c_1,c_2$ are determined by the boundary coefficients. \\
    We know that $\alpha(0)=1$ and $\alpha(d)=1$. Simplifying $1-4(pq)=1-4p(1-p)=1-4p+4p^2=(1-2p)^2$ then $\beta_{1,2}=\frac{1\pm(1-2p)}{2p}$. So $\beta_1=\frac{2-2p}{2p}=\frac{q}{p}, \beta_2=1$ (still assuming $0<p<1$). So the final form is $\alpha(j)=c_1(\frac{p}{q})^j+c_2$. \\
    Using the boundary conditions $\alpha(0)=1=c_1+c_2$ and $\alpha(d)=0=c_1(\frac{p}{q})^d+c_2$. This means that $c_2=1-c_1$. Then $1=c_1(\frac{p}{q})^d+1-c_1$ or $c_1=c_1(\frac{p}{q})^d$. This is only possible if $c_1=0$. So we get $\alpha(j)=1$.
\end{enumerate}
\end{center}
\end{document}