\documentclass[14pt,letterpaper]{extarticle}
\usepackage[utf8]{inputenc}

% Set margins
% https://tex.stackexchange.com/a/327136
% this should replace use of the 'fullpage' package
\usepackage[
    includehead,
    margin=2.5cm,
    bmargin=2cm,
]{geometry}

\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{verbatim} % This works better than comment for some reason
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{gensymb}
\usepackage{enumitem}
\usepackage{xfrac} % Fractions - https://tex.stackexchange.com/a/230973
\usepackage{xcolor}
% The magic latex drawing thing
\usepackage{tikz}
\usetikzlibrary{shapes}
\usepackage[en-US,showdow]{datetime2}

% Colored hyperlinks:
% https://www.overleaf.com/learn/latex/Hyperlinks
\hypersetup{
    colorlinks=true,
    urlcolor=blue
}

% Setup nice headings
% https://www.overleaf.com/learn/latex/Headers_and_footers
\usepackage{fancyhdr}
\usepackage{titling} % Save title for use in fancyhdr

% Shortcuts
\newcommand*{\bP}{\mathbf{P}}



% Use black tombstone for QED
\renewcommand{\qedsymbol}{$\blacksquare$}

% Important sets
\newcommand*{\C}{\mathbb{C}}
\newcommand*{\R}{\mathbb{R}}
\newcommand*{\Q}{\mathbb{Q}}
\newcommand*{\Z}{\mathbb{Z}}

%
% symbols
%

% name by meaning, not by character
\newcommand{\grad}{\nabla}


% New command for 'partial QED' using white tombstone
\newcommand*{\partialqed}{$\hfill \square$ \\}

\newcommand*{\todo}{{\color{red}\textbf{TODO:}} }

% Increased spaces spacing
\usepackage{setspace} \onehalfspacing

\title{Math 468 - Homework 1}
\author{Nicholas Schlabach (nickninja@arizona.edu)}
\DTMsavedate{titledate}{2026-02-04}
\date{\DTMUsedate{titledate}}

%% Copied from https://tex.stackexchange.com/a/2238
%% Modified by Techcable 9/19/21 to use '[...]' instead of '(...)'
\newenvironment{amatrix}[1]{%
  \left[\begin{array}{@{}*{#1}{c}|c@{}}
}{%
  \end{array}\right]
}

\begin{document}
\pagestyle{fancy}
\fancyhead[L]{Nicholas Schlabach}
\fancyhead[C]{\hspace{5mm}\Big(\thetitle{}\Big)}
\fancyhead[R]{{\DTMsetdatestyle{mmddyy}\DTMusedate{titledate}}, Page \thepage}
\fancyfoot{} % no footer
\maketitle
% maketitle automatically sets pagestyle=plain, override this
\thispagestyle{fancy}


\begin{center}
\begin{enumerate}
\item
    \begin{enumerate}
    \item
        This case has the following transition matrix:
        \[ \bP=\begin{bmatrix}
            1-p & p \\
            0 & 1 
        \end{bmatrix} \]
        Since $\bP(1,0)=q=0$, then $\rho_{10}=0$ and state $s_1$ does not lead to state $s_0$. This means $s_0$ and $s_1$ do not communicate and so fall into separate communication classes, making this a reducible markov chain. Since $\bP(1,1)=1$ then $s_1$ is absorbing and  $\rho_{11}=1$. This means that $s_1$ is also recurrent. Since $P(0,1)>0\implies \rho_{01}>1$ and $1$ does not lead back to $0$, then state $s_0$ is transient. The communication class $\{s_1\}$ is closed but $\{s_0\}$ is not closed.
    \item
        In the case $p=q=0$,  the transition matrix $\bP$ is the identity matrix. Both states $s_0, s_1$ are absorbing meaning they are both recurrent. Neither state leads to the other so there are two communication classes and the system is irreducible. Both communication classes are closed. \\
        In the case $p=q=1$, both states communicate $s_0 \leftrightarrow s_1$ and so there is a single cloased communication class containing both states. In fact, because $P(0,1)=1=P(1,0)$ then both states are recurrent.
    \end{enumerate}
\setcounter{enumi}{2}
\item
    \begin{enumerate}
    \item
        BWOC, assume that the states are not all distinct, so that $x_m=x_k$ for some $m\neq k$.
        WLOG, assme that $m<k$.
        Because $P(x_i,x_{i+1})>0$ for all $0< i<n_0$, then the following series of state transitions has nonzero probability: $xx_1\dots x_mx_{k+1}\dots x_{n-1}y$.
        This chain has length $n^*:=n_0-(k-m)<n_0$. This means that $\bP^{n^*}(x,y)>0$ with $n^*<n_0$, contradicting the  the definition of $n_0$ being the samllest integer satisfying that property.
    \item
        Because of (a), the sequence of transitions $xx_1,\dots, x_{n-1}y$ has no duplicates states and so the corresponding set $\{x,x_1,\dots,x_{n-1}y\}$ has cardinality $n_0+1$.
        However, since there are only $d$ states in total, than $n_0+1\le d$. This is equivalent to the desired result.
    \item
        By definition, $n_0$ is the smallest number such that a sequence of state transitions from $x$ to $y$ exists taking $n_0$ steps.
        But this sequence of states with length $n_0$ having nonzero probability is equivalent to saying that $P_x(T_y=n_0)>0$. We know from part (b) that $n_0\le d-1$, meaning that $P_x(T_y\le d-1)>0$.
    \end{enumerate}
\item
    \begin{enumerate}
    \item
        In the case of the random walk on the unit circle, all states fall into a single communication class, making the system irreducible. This is true regardless of whether $p=1$ or $p=0$ due to the effect of wraparound.
    \item
        
        Label the states $s_1s_2\dots s_N$. Assume that $N\ge 2$ or else we get a degenerate case with a single state. \\In the case where $p=1$, then each of the states $s_1,s_2,\dots s_{N-2}$ fall into their own communication class. The states $s_N \leftrightarrow s_{N-1}$ communicate with each other as $s_N$ is reflecting. This means there are $N-1$ commucation classes and the system is reducible unless $N=2$. The same thing happens in the $q=1$ case due to symmetry. \\
        \vspace{3mm}
        When $0<p<1$, then all the states communicate with each other and so they all fall into a single communication class. This means that the system is irreducible. 
    \end{enumerate}
\end{enumerate}
\end{center}
\end{document}